---
title: "Descripte analysis"
author: "Lucas Avezum"
date: "June 17, 2019"
output:
  html_document: 
    toc: true
  pdf_document:default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)                                                           
library(data.table)
library(lubridate)
library(lfe)
library(stargazer)
library(ineq)
library(DescTools)
library(gglorenz)
library(ggplot2)
library(psych)
library(reporttools)
library(ggridges)
library(splitstackshape)
library(viridis)
library(Hmisc)
library(scales)

```
```{r load data, include=FALSE}
rm(list = ls())   
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankScope.Rda")
data.pillar3 <- data.pillar3 %>%
  mutate(Method2 = as.factor(ifelse(Method %in% c("IRB"), "IRB",
                                                     sapply(Method,str_sub,start= 2,end=4)))) 
bankscope <- bankscope %>%
  mutate(sample = as.factor(ifelse(bvdid %in% data.pillar3$bvdid,"Yes","No")),
         mean.ROA= ave(ROA.using.P.L.before.tax....,bvdid,FUN = function(x) mean(x, na.rm = TRUE)),
         sd.ROA= ave(ROA.using.P.L.before.tax....,bvdid,FUN = function(x) sd(x, na.rm = TRUE)),
         mean.CAR = ave(Equity...Total.assets....,bvdid,FUN = function(x) mean(x, na.rm = TRUE)))%>%
mutate('Z-score'=(mean.ROA+mean.CAR)/sd.ROA)
```

```{r expand data, include=FALSE}

expanded.pillar3 <- data.pillar3 %>%
  filter(Method2 %in% c("IRB")) %>%
  select(Bank,Country,Method,`Portfolio - level 1`,EAD,LGD,PD,year)%>%
  mutate(EL=round(EAD*LGD/1000),EAD=round(EAD/1000),PD=PD*100)


expanded.list <- list()
for(i in c("EAD","EL")) {
expanded.list[[i]] <- expanded.pillar3  %>%
  na.omit(cols=c(i))%>%
  expandRows(i) %>%
  mutate(freq_type = as.factor(i))
}
expanded.pillar3 <-  rbindlist(expanded.list, use.names=TRUE,fill = TRUE)

```

## Data

In this report I briefly describe the data collected so far and provide some descriptive analysis of two measures of capital savings. \newline

Table 1 reports summary statistics for a few variables from Bankscope. The level division means if the bank's Pillar-III reports information has been collected. The p-values from Kruskal test reject the hypothesis that the two groups are draw from the same distribution. The banks collected tend to have higher leverage (both risk-weighted or not), be safer (higher Z-score and lower NPL to gross loans ratio), and higher return on equity.      

```{r sample vs bankscope, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
options(xtable.comment = FALSE)
vars <-bankscope%>%
  select(ROA = ROA.using.P.L.before.tax....,
         ROE =ROE.using.P.L.before.tax....,
         `NPL/gross loans` = Non.Perf..Loans...Gross.Loans....,
         `Total capital ratio` = Total.Capital.Ratio....,
         `Equity/total assets` = Equity...Total.assets....,
         `Z-score`) %>%
  mutate_all(Winsorize,na.rm = TRUE, probs = c(0.025, 0.975))


tableContinuous(vars = vars, group = bankscope$sample, longtable = FALSE,prec = 1, 
                cap = "Summary statistics", lab = "tab: descr stat", print.pval = "kruskal",
                stats = c("n","mean", "s", "min","median", "max"))

```
```{r  PD distribution by bank,fig.height = 7, fig.width = 7,eval=FALSE, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(Bank, PD,FUN = median, na.rm=TRUE))) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2,
                       scale = 2 , from = 0, to = 2.5) +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    title = 'PD distribution by bank',
    subtitle = 'Data: Pillar-III reports'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + theme(axis.title.y = element_blank())
```

Figure 1 plots the PD distribution collected by bank.

```{r  PD distribution by bank and country,fig.height = 7, fig.width = 8, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(Bank, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 2.5) +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    title = 'Figure 1: PD distribution by bank',
    subtitle = 'Ordered by median PD'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + theme(axis.title.y = element_blank())
```


## Basel formulas and capital savings

$K=\sum_{i}EAD_i\times LGD_i\times\bigg[\Phi\bigg(\frac{\Phi^{-1}(PD_i)+\sqrt{\rho_i}\Phi^{-1}(0.999)}{\sqrt{1-\rho_i}}\bigg)-PD_i\bigg]$ \newline

For wholesale portfolio:  \newline
$K_{\text{wholesale}}=K\times \frac{1+(M-2.5)b(PD_i))}{1-1.5b(PD_i)}$ \newline
with\newline
$\rho_i = 0.12\times\bigg(\frac{1-e^{-50\times PD_i}}{1-e^{-50}}+0.24\times\bigg(1-\frac{1-e^{-50\times PD_i}}{1-e^{-50}}\bigg)\bigg)$\newline

and $b(PD_i)=(0.11852-0.05478\times log(PD_i))^2$\newline

For retail portfolios the formula is equal to $K$ with different formulas for the correlation coefficient:  \newline

Real estate: $\rho_{i}=0.15$\newline

Qualifying revolving: $\rho_{i}=0.04$ \newline

Other retail: $0.03\times\bigg(\frac{1-e^{-35\times PD_i}}{1-e^{-35}}+0.16\times\bigg(1-\frac{1-e^{-35\times PD_i}}{1-e^{-35}}\bigg)\bigg)$\newline


Using these formulas I computed two measures of capital requirement: $K_{distribution}$ using the distribution of $PD_i$ (as it is required by the Basel accords), and $K_{average}$ using $\widetilde{PD}=\sum_iEAD_iLGD_iPD_i/\sum_iEAD_iLGD_i$. The percentage change $K_{distribution}/K_{average}-1$ is called capital savings and is a measure of the amount of capital saved due to the concavity in the formula for $K$.\newline

```{r mapping formulas, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

mapping_wholesale <- function(PD,LGD,M,alpha,portfolio){
  R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
  b  <- (0.11852-0.05478*log(PD))^2
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
  return(data)
}

mapping_retail <- function(PD,LGD,R,alpha,portfolio){
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,R,portfolio)
  return(data)
}


mapping_other_retail <- function(PD,LGD,alpha,portfolio){
  R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,R,portfolio)
  return(data)
}
```

```{r savings calculation, echo=FALSE, results='asis', warning=FALSE, message=FALSE}


data.pillar3   <- data.pillar3 %>%
  filter(Method2 == "IRB",PD<1)%>%
  mutate(K_original=RW*EAD*0.08,
         PD_w=PD*EAD*LGD,
         EAD_w=LGD*EAD) %>%
  mutate(EAD_sum=ave(EAD_w,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`,FUN = function(x) sum(x,na.rm = TRUE)),
         PD_w_sum=ave(PD_w,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`,FUN = function(x) sum(x,na.rm = TRUE)))%>%
  mutate(PD_w_avg = PD_w_sum/EAD_sum)

for (i in 1:nrow(data.pillar3)){
  if (data.pillar3[i,9]!="Wholesale"){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_wholesale(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],2.5,0.999,"")["K"]
  
  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_wholesale(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],2.5,0.999,"")["K"]
}

for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (!data.pillar3[i,10]%in%c("Real estate")){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.15,0.999,"")["K"]
  
  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.15,0.999,"")["K"]
  
}

for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (!data.pillar3[i,10]%in%c("Qualifying revolving")){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.04,0.999,"")["K"]
  
  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.04,0.999,"")["K"]
  
}
for (i in 1:nrow(data.pillar3)){
  if (is.na(data.pillar3[i,10])){
    next
  }
  if (data.pillar3[i,10]!="Other retail"){
    next
  }
  data.pillar3[i,"K_calculated"] <- data.pillar3[i,"EAD"]*mapping_other_retail(data.pillar3[i,"PD"],data.pillar3[i,"LGD"],0.999,"")["K"]
  
  data.pillar3[i,"K_counterfactual"] <- data.pillar3[i,"EAD"]*mapping_other_retail(data.pillar3[i,"PD_w_avg"],data.pillar3[i,"LGD"],0.999,"")["K"] 
  
}

data.reduced <- data.pillar3 %>% group_by(Bank,Country,year) %>%
  summarise_if(is.numeric,funs(sum),na.rm=TRUE) %>%
  mutate(error = (K_calculated - K_original)/K_original,
         savings = (K_calculated-K_counterfactual)/K_counterfactual) %>%
  ungroup()


```

Figure 2 plots the calculated capital savings by country. Figure 3 plots the calculated capital savings by country. Figure 4 plots the evolution of calculated capital savings in time. \newline


```{r savings by bank, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.reduced %>%
 ggplot(
            mapping = aes(x = reorder(Bank, savings,FUN = median, na.rm=TRUE),
                          y = savings,
                          color = Country,
                          fill = Country)) +
geom_boxplot(alpha = 0.5) +
    labs(x=NULL) +
    coord_flip() +
  labs(
       y =  "Capital savings",
         title = "Figure 2: Capital savings by bank") +
  
  theme_bw()

```



```{r savings by country, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.reduced %>%
 ggplot(
            mapping = aes(x = reorder(Country, savings,FUN = median, na.rm=TRUE),
                          y = savings)) +
geom_boxplot() +
    labs(x=NULL) +
    coord_flip() +
  labs(
       y =  "Capital savings",
         title = "Figure 3: Capital savings by country") +
  theme_bw()

```



```{r savings by year, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = na.rm),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

data.reduced %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time, y=savings)) +
 geom_smooth(stat = 'summary', fun.data = mean_cl_quantile)+
labs(y =  "Capital savings",
     x = "Year",
     title = 'Figure 3: Average capital savings by year',
     subtitle = 'First and third quartile shown') +
  theme_bw()
```

## Basel formulas and Gini coefficient

As a second measure of the $PD$ distribution dispersion I compute Gini coefficients. Figure 4 shows the Lorenz curve of the $PD$ distributions  by country. Figure 5 shows the Lorenz curve of the $PD$ distributions by year. Figure 6 shows the Lorenz curve of the $PD$ distributions by Bank (names omitted for visibility). Finally Figure 7 plots the measure of capital savings and the Gini coefficients. \newline    

```{r lorenz curve by country, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = Country)) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
   labs(
       y =  "Lorenz curve",
         title = "Figure 4: Lorenz curve by country") +
theme_minimal()

```



```{r lorenz curve by year, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = as.factor(year))) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
  labs(color = "",
       y =  "Lorenz curve",
         title = "Figure 5: Lorenz curve by year") +
theme_minimal()
```



```{r lorenz curve by bank, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
expanded.pillar3 %>% 
  filter( !is.na(PD), freq_type%in%c("EL") ) %>% 
ggplot( aes(x=PD, color = Bank)) +
stat_lorenz() +
coord_fixed() +
geom_abline(linetype = "dashed") +
  theme_minimal()+
  labs(
       y =  "Lorenz curve",
         title = "Figure 6: Lorenz curve by bank") +
theme(legend.position = "none")
```



```{r gini savings total, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
data.gini <- data.pillar3 %>% 
  group_by(Country,Bank,year,`Portfolio - level 1`,`Portfolio - level 2`) %>%
  mutate(weight=EAD*LGD)%>%
  filter(weight>0, `Portfolio - level 1` %in% c("Wholesale","Retail"))%>%
  dplyr::summarize(gini = DescTools::Gini(PD,weight),
            K_calculated=sum(K_calculated),
            K_counterfactual=sum(K_counterfactual)) %>%
  mutate(savings = (K_counterfactual-K_calculated)/K_counterfactual)

p <- ggplot(data = data.gini,
            mapping = aes(x = gini, y = savings, color=`Portfolio - level 1`))

 p + geom_point(alpha = 0.2)+
theme_minimal()+
   labs( title = "Figure 7: Capital savings and Gini coefficient",
         subtitle = 'Each point is a bank-year-portfolio observation',
         y =  "Inverted capital savings",
         x = "Gini coefficient",
         color = "")
```




```{r summary,eval=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
bankscope %>% 
  subset(select = c("ROA.using.P.L.before.tax....","ROE.using.P.L.before.tax....","Non.Perf..Loans...Gross.Loans....","Total.Capital.Ratio....","Equity...Total.assets....","Employees","Total.assets","year","country","bvdid")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "footnotesize",
          omit.summary.stat = c("p25","p75"))
``` 



